---
title: Tutorials in Positron
author: David Kane
tutorial:
  id: tutorials-in-positron
output:
  learnr::tutorial:
    progressive: yes
    allow_skip: yes
runtime: shiny_prerendered
description: Tutorial for introducing students to tutorials when using Positron.
---

```{r setup, include = FALSE}
library(learnr)
library(tutorial.helpers)
library(remotes)

knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(out.width = '90%')
options(tutorial.exercise.timelimit = 60, 
        tutorial.storage = "local") 
```

```{r copy-code-chunk, child = system.file("child_documents/copy_button.Rmd", package = "tutorial.helpers")}
```

## Introduction
### 

This tutorial is an introduction, for students using [Positron](https://positron.posit.co/), to the general use of R tutorials created with the [**learnr**](https://rstudio.github.io/learnr/) package. It is meant for beginning students who have never completed a tutorial.

### 

<!-- DK: Update computer set up link. -->
<!-- DK: Add new document discussing Workbench. -->

You may complete a tutorial either on your laptop or on a cloud computing platform like [Posit Workbench](https://posit.co/products/enterprise/workbench/). (See our discussion about [setting up your own computer](https://ppbds.github.io/primer/getting-started.html).) Either way, you will download a copy of your answers to the Downloads folder on your computer once you are done with a tutorial. Then, you will submit your answers to your instructor and/or keep a copy for yourself.

### 

Most tutorials begin by asking for your your name and email. (Your instructor might also request that you fill in the "ID" field so that your work can be matched to a grading database. If not, leave the "ID" field blank.) Click "Continue" and then fill in your name and email. Click the **"Submit Answer"** button each time. If you make a mistake, you can click **"Try Again"** and, uh, try again. 


```{r info-section, child = system.file("child_documents/info_section.Rmd", package = "tutorial.helpers")}
```

## Exercises 
### 

In each tutorial, there will be a set of exercises for you to complete. For non-coding questions, you write your answer and then hit the "Submit Answer" button, as you did above.

### 

Clicking **"Start Over"** --- look for small letters at the bottom of the left-hand side menu --- deletes all your work for a tutorial and allows you to begin the entire tutorial again. (If you are working on your tutorial in the Viewer, the screen might not be wide enough to show the left-hand side menu. If so, just expand the Viewer until it appears.)

<!-- DK: Also, or better, we can just click on the tutorial name in the top left and Start Over appears at the bottom of the resulting menu. -->

<!-- DK: Test to see if you can just pull out the whole pane into a separate window. -->

### Exercise 1

Let's practice some simple math. Type `4 + 5` into the box below. Hit **"Run Code"** to see your results. Clicking **"Hint"** gives you a hint about what to type. 

```{r exercises-1, exercise = TRUE}

```

```{r exercises-1-hint-1, eval = FALSE}
4 + 5
```

### 

Once you click "Run Code," the answer `[1] 9` should appear below the code you entered. This is the same behavior you would see if you ran this code in the Console, which is where R is run, in a tab beneath the Welcome page.

### 

There are two types of answers for tutorial questions: prose and code. With prose answers, you will use the "Submit Answer" button, as you did with your name and email. With code answers, you use the "Run Code" button, as you did for this question.

### 

If you don't "Run Code" (or "Submit Answer") your response, nothing is recorded.

### Exercise 2

Let's use variables to perform the same calculation. Below, we have already assigned 5 to `x` and 4 to `y`. On the line below those two provided lines of code, add the two variables by typing `x + y`. Hit "Run Code" to see your results. 

```{r exercises-2, exercise = TRUE}
x <- 5
y <- 4

```

```{r exercises-2-hint-1, eval = FALSE}
x <- 5
y <- 4
x + y
```

### 

Once you click "Run Code", the answer `[1] 9` should appear below the code you entered. 

### Exercise 3

We will often need to copy code from our previous exercises to continue working on a multi-stage problem. To make this easier, we provide a "Copy previous code" button. Below the following box, click on the "Copy previous code" button. 

```{r exercises-3, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r exercises-3-hint-1, eval = FALSE}
x <- ...
y <- 10
x + ...
```

### 

This should paste all work from the previous exercise into the above exercise code area. Note that this will only copy code which has already been run in the previous exercise.

### 

Change the values to `x <- 6` and `y <- 10`. Hit "Run Code." You should get the answer `[1] 16`.

### 

Note that the assignment symbol, `<-`, is a less-than-sign followed by a dash. We often pronounce the symbol as "gets." So, we might say, when reading this code, that "x gets 6."

### 

Did you remember to hit the "Run Code" button above? Lots of people forget! You must hit the button for your answer to be recorded.

### 

If you used the hint, note how we did not provide the entire answer. This is our standard approach with hints.




### Exercise 4

<!-- DK: Seems like it should be possible to restart the R session. But the various buttons don't seem to work very well, at least when you are in a tutorial. Try this again in a month. In the meantime, this radical close-Positron approach is OK. -->

Read all of the following instructions before continuing.

* Close and then restart Positron. (The purpose of this exercise is to demonstrate that your work is saved automatically.)

* Restart the tutorial. The most common approach is to run this command at the Console:

```
learnr::run_tutorial("tutorials-in-positron", package = "tutorial.helpers")
```

Now that you have read all the instructions, go ahead and follow them.

### 

Note how all your answers to previous questions have been saved. It is not a problem to quit a tutorial. In fact, if you ignore a tutorial for awhile, it will quit automatically.


### Exercise 5

We often ask you to execute an R command in the Console, a pane on the lower left portion of your Positron screen. Type `version[['version.string']]` into the Console and hit the return/enter key.

**Do not just type the command into the answer box!** You must paste the command into the Console and then execute the command.

Copy/paste the command **and** the return value into this box.

```{r exercises-5}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 3)
```

### 

Your answer should look something like:

````
> version[['version.string']]
[1] "R version 4.4.2 (2024-10-31)"
> 
````

Your version of R may be more recent, but it should not be older. It is OK if your formatting differs from ours. Exactness does not matter. What matters is that you prove that you have executed the command as instructed.

### 

Tutorial workflow will often look like this. We tell you to run a command **in the Console**. You run it. We then ask you to copy/paste, from the Console, both the command itself and the response returned. Because we ask you to **c**opy/**p**aste the **c**ommand/**r**esponse so often, we usually shorten this instruction to CP/CR.

### Exercise 6

Positron and your R installation have some default settings which are not good for new users. Run this command at the Console to change them:

`
tutorial.helpers::set_positron_settings()
`

CP/CR.

```{r exercises-6}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 8)
```

### 

The output from this function will report the changes, if any, which were made. To learn more, look up the help page by running `?tutorial.helpers::set_positron_settings` at the Console.

### 

You can generally call functions by just giving their name along with the ending parentheses, as with `set_positron_settings()`. However, it can be safer to preface the function name with the name of the package in which the function is found, separated by two colons. Since `set_positron_settings()` is in the **tutorial.helpers** package, we can also call it using the double colon notation: `tutorial.helpers::set_positron_settings()`, as we do above.

### Exercise 7

Tutorials can include written responses. You have already seen several examples. Sometimes those written answers are just copies of the R commands and their results. Other times, we will ask you to write one or more sentences of prose. 

Copy and paste everything from `Exercise 7` above through the end of this sentence into the answer box below. That is, you are copying from this tutorial, not from the Console. Press "Submit Answer."

```{r exercises-7}
question_text(NULL,
    message = "We will sometimes provide an answer to written questions. Other times, like here, we won't. In the former case, we will not allow you to resubmit your answer. Don't stress! Most instructors grade tutorials on pass/fail basis, so, as long as you make an honest effort, you will do fine.",          
    answer(NULL, 
           correct = TRUE),
    allow_retry = FALSE,
    incorrect = NULL,
    rows = 6)
```

### Exercise 8

`show_file()` is another commonly-used function from the **tutorial.helpers** package. It makes it easy for you to share the contents of another file with your instructor, especially a file which you have edited. 

From the Console, run this command:

```
tutorial.helpers::show_file(file.path(R.home(), "COPYING"), end = 7)
```

CP/CR.

```{r exercises-8}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 3)
```

<!-- I really want a test case to ensure that this works on all OSs. -->

```{r exercises-8-test, include = FALSE}
tutorial.helpers::show_file(file.path(R.home(), "COPYING"), end = 7)
```

### 

Your answer should look something like this:

````
> tutorial.helpers::show_file(file.path(R.home(), "COPYING"), end = 7)
		    GNU GENERAL PUBLIC LICENSE
		       Version 2, June 1991

 Copyright (C) 1989, 1991 Free Software Foundation, Inc.
                       51 Franklin St, Fifth Floor, Boston, MA  02110-1301  USA
 Everyone is permitted to copy and distribute verbatim copies
 of this license document, but changing it is not allowed.
> 
````

### 

For now, don't worry about what `R.home()` and `file.path()` are doing, although you can read about them by running `?R.home` and `?file.path` at the Console. Those details are less important than seeing the intended usage of `show_file()`. 


## Your answers
### 

At the end of every tutorial, you will save your work (and submit it, at the discretion of your instructor). We will do this in the next topic, labeled "Download answers." Follow the instructions there. 

Make sure to answer the question about how many **minutes** you spent on the tutorial and to click "Submit Answer."

After hitting the "Download" button, you will be prompted to save a file named `getting-started_answers.html`. This file will be saved in the default Downloads folder on your computer, even if you completed the tutorial on a cloud computing platform. 

Unless told otherwise by your instructor, you should use the default file name, which will usually be of the form `name-of-tutorial` + `_answers` + `.html`. 

```{r download-answers, child = system.file("child_documents/download_answers.Rmd", package = "tutorial.helpers")}
```

## Summary
### 

You now understand how tutorials function and how to download a copy of your answers. Good luck with your data science journey!
