---
title: Introduction to Python
author: Satvika Upperla
tutorial:
  id: introduction-to-python
output:
  learnr::tutorial:
    progressive: yes
    allow_skip: yes
runtime: shiny_prerendered
description: Tutorial for introducing students to the Python language.
---

```{r setup, include = FALSE}
library(learnr)
library(tutorial.helpers)
library(knitr)
library(tidyverse)

knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(out.width = '90%')
options(tutorial.exercise.timelimit = 60, 
        tutorial.storage = "local") 
```

```{r copy-code-chunk, child = system.file("child_documents/copy_button.Rmd", package = "tutorial.helpers")}
```

```{r info-section, child = system.file("child_documents/info_section.Rmd", package = "tutorial.helpers")}
```

### Introduction
##

This is tutorial introduces you to Python from a data science standpoint. We will explore Polars, a python data fram for data processing, that is more fast and efficient at handling large datasets over Pandas and Numpy two other pYthon data frames.


### Setting up Python on Your Console
##

Even if you already have python installed on you setup, its recommended you still follow this, unless you have access python striaght from your console pane in positron.

1. Run Python in Terminal, this will pull up Microsfot Store's version of Python 3.17.3 or later. Install. Once the File downloads, run the .exe file and complete setup. **Remember to note the file path it installs Python in**

2. Click the icon in the lower left corner of Positron (Settings) and go to settings. Press Interpreter in the little arrow bar menu. Scroll down to Python. 

3. In the Python section there are two spots you need to edit. One is the Default interpreter path. The other is Interpreters include: where you can press an add button to set you own interepreter.

4. Add the path to your python installation in default interpreter path. Here is an example `C:\Users\USatv\AppData\Local\Programs\Python\Python311\python.exe`. Add this same path to the other section as well (Interpreter include:) using the add button. In general, your path will always be "C:Users\" and then whatever user file your laptop will have made for your person, and then "\AppData\Local\Programs\Python\Python311\python.exe". The Python311 may differ if you have any later version of Python past the 3.13 line.

5. Restart the Positron window. This can be done by searching up `> Developer: Reload Window` in the search bar up above.

```{r download-answers, child = system.file("child_documents/download_answers.Rmd", package = "tutorial.helpers")}
```
