---
title: 'Test Tutorial'
author: 'David Kane'
tutorial:
  id: test-tutorial
output:
  learnr::tutorial:
    progressive: yes
    allow_skip: yes
runtime: shiny_prerendered
description: 'Tutorial for use in testing'
---

```{r setup, include = FALSE}
library(learnr)
library(tutorial.helpers)
knitr::opts_chunk$set(echo = FALSE)
options(tutorial.exercise.timelimit = 60, 
        tutorial.storage = "local") 

```

```{r copy-code-chunk, child = system.file("child_documents/copy_button.Rmd", package = "tutorial.helpers")}
```

```{r info-section, child = system.file("child_documents/info_section.Rmd", package = "tutorial.helpers")}
```

## Introduction
### 

This tutorial covers [Chapter 25: Web scraping](https://r4ds.hadley.nz/webscraping.html) from [*R for Data Science (2e)*](https://r4ds.hadley.nz/) by Hadley Wickham, Mine Çetinkaya-Rundel, and Garrett Grolemund.


## Web scaping basics
### 


This section covers the basics of web scraping.

### Exercise 1

The key package for web scraping with R is [**rvest**](https://rvest.tidyverse.org/). Load the **rvest** library.

```{r web-scaping-basics-1, exercise = TRUE}

```

```{r web-scaping-basics-1-hint-1, eval = FALSE}
library(rvest)
```

### 

The "Web Scraping 101" [article](https://rvest.tidyverse.org/articles/rvest.html) from the **rvest** [homepage](https://rvest.tidyverse.org/) provides useful background. 

### Exercise 2

Look up the help page for the **rvest** package using the `help()` function. Don't forget to use the `package` argument.

```{r web-scaping-basics-2, exercise = TRUE}

```

```{r web-scaping-basics-2-hint-1, eval = FALSE}
help(package = "...")
```

### 

The help page for a package provides a listing of all the functions in a package. Note that there are both `html_text()` and `html_text2()` functions. You should always use `html_text2()` unless you have a good reason not to.

### Exercise 3

Click on the link to the DESCRIPTION file which appeared when you looked up the help page for **rvest**. Copy the the first two lines (Package and Title) below.

```{r web-scaping-basics-3}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 3)
```

### 

As the DESCRIPTION file mentions, **rvest** is mostly supplying wrappers for the more lower level (but harder to use) **xml2** and **httr** packages. (As always, note how we, and everyone else uses the words "package" and "library" interchangeably.)



## Summary
### 

This tutorial covered [Chapter 25: Web scraping](https://r4ds.hadley.nz/webscraping.html) from [*R for Data Science (2e)*](https://r4ds.hadley.nz/) by Hadley Wickham, Mine Çetinkaya-Rundel, and Garrett Grolemund.



```{r download-answers, child = system.file("child_documents/download_answers.Rmd", package = "tutorial.helpers")}
```
